---
title: "Global Helpers"
author: "Erwan Guyomarch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
header-includes:
   - \usepackage{babel}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
# library(DT)
# require(devtools)
# install_github('ramnathv/rCharts')
```


## Header

You can import latex packages adding this chunk in the .Rmd header:
```

```




## Useful functions

```{r}
# convert the numbers to formatted strings
comma <- function(x) format(x, digits = 2, big.mark = ",")

# transform an excel integer-formatted date to a r date
get_date <- function(x, format = '%m.%y') {
  ifelse(!is.na(as.numeric(x)), format(as.Date(as.numeric(x), origin="1899-12-30"), format), x)
}

# copy a small data.frame to the clipboard
tc <- function(x, row.names=FALSE, col.names=TRUE, ...) {
  write.table(x, "clipboard", sep="\t", row.names=row.names, col.names=col.names, ...)
}
```




## Import

Base R ([doc](https://www.rdocumentation.org/packages/utils/versions/3.6.1/topics/read.table)):
```{r}
# import from clipboard
my_data <- read.table(file = "clipboard")
my_data <- read.table(file = "clipboard", sep = "\t", header=TRUE)

# from a CSV
my_data <- read.table("c:/mydata.csv", header=TRUE, sep=",", row.names="id")
```

Using *readr*:
```{r}
read_csv("challenge-2.csv")
```

Using *readxl*:
```{r}
library(readxl)
read_excel("List_activities_diab.xlsx")
```

Using *xlsx* (`.xls` and `.xlsx`):
```{r}
file <- system.file("tests", "test_import.xlsx", package = "xlsx")
# Preserves the type of data
read.xlsx(file, sheetIndex, header=TRUE, colClasses=NA)
# Faster
read.xlsx2(file, sheetIndex, header=TRUE, colClasses="character")
```



## Export

### CSV Export

Base R:
```{r}
write.table(data, file = "MyData.csv", row.names=FALSE, na="", col.names=FALSE, sep=",")
```

Using *readr*:
```{r}
write_csv(data, "challenge-2.csv")
```

Using *xlsx*:
```{r}
write.xlsx(data, "filename.xls", sheetName="Sheet1", col.names=TRUE, row.names=TRUE, append=FALSE)
write.xlsx(data, "filename.xlsx")
# faster
write.xlsx2(data, "filename.xlsx")
```


### Complex Excel export

You can create a file excel to export a large data set. 
To create an Excel file:

- First have to create a *workbook* using `create_wb()`.

```{r}
library(openxlsx)
wb <- createWorkbook(title = "DataExport", creator = "Erwan Guyomarch")
```

- Then you have to create a new sheet using `wb_sheet()`. You can modify a row using `wd_row()`.

```{r}
addWorksheet(wb, "NewSheet")
writeData(wb, "NewSheet", data) 
```

- Finally you can create the Excel file in the working folder using `save_wb()`:

```{r}
file_name <- "filename.xlsx"
local_path <- dirname(rstudioapi::getSourceEditorContext()$path)
saveWorkbook(wb, paste(local_path, name, sep="/"), TRUE)
```

### Modification of the columns

```{r}

```




## Other

- LaTeX array generator: <https://www.tablesgenerator.com/#>
  - Add \begin{table}[ht]
