---
title: "Global Helpers"
subtitle: "R Cheat Sheet"
author: "Erwan Guyomarch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
# library(DT)
# require(devtools)
# install_github('ramnathv/rCharts')
```


## Setup

### Header

Basic:
```
---
title: "MyTitle"
subtitle: "MySubtitle"
author: "MyName"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```

Pdf:
```
output: 
  pdf_document:
    toc: TRUE
    toc_depth: 4
    number_sections: FALSE
classoption: twocolumn
always_allow_html: yes
latex_engine: xelatex
geometry: left = 0.6cm, right = 0.6cm, top = 1.5cm, bottom = 1.5cm
mainfont: Arial
header-includes:
  - \usepackage{booktabs}
  - \setlength{\columnsep}{0.5cm}
```

Reveal.js:
```
output: 
  revealjs::revealjs_presentation:
    css: reveal_css_fix.css
    self_contained: false
    reveal_plugins: ["notes", "zoom"]
    theme: night # black / white / league / sky / beige / simple / serif / blood / night / moon / solarized
    center: true
    transition: slide
    reveal_options:
      controls: false
      slideNumber: true
      hash: true
      history: true
      defaultTiming: 120
      mouseWheel: true
```



### Chunks

The chunk options are:

- `eval = TRUE`: the code chunk is interpreted
- `include = TRUE`: 
- `echo = TRUE`: 
- `warning = TRUE`: 
- `error = FALSE`: 
- `message = TRUE`: 
- `tidy = FALSE`: 
- `results = "markup/hide/asis/hold"`: 
- `cache = FALSE`: 
- `comment = "##"`: 
- `fig.width/fig.height = 7`: 
- `fig.align = "left/right/center"`: 


## Useful Shortcuts

```{r}
# list the files in the working directory
list.files()

# delete all the stored variables
rm(list = ls())
```



## Useful Functions

```{r}
# convert the numbers to formatted strings
comma <- function(x) format(x, digits = 2, big.mark = ",")

# transform an excel integer-formatted date to a r date
get_date <- function(x, format = '%m.%y') {
  ifelse(!is.na(as.numeric(x)), format(as.Date(as.numeric(x), origin="1899-12-30"), format), x)
}

# copy a small data.frame to the clipboard
tc <- function(x, row.names=FALSE, col.names=TRUE, ...) {
  write.table(x, "clipboard", sep="\t", row.names=row.names, col.names=col.names, ...)
}
```




## Import Data

### Clipboard

Base R ([doc](https://www.rdocumentation.org/packages/utils/versions/3.6.1/topics/read.table)) (*.csv* and *Clipboard*):
```{r}
# import from clipboard
my_data <- read.table(file = "clipboard")
my_data <- read.table(file = "clipboard", sep = "\t", header=TRUE)
```

### CSV

Base R:
```{r}
my_data <- read.table("c:/mydata.csv", header=TRUE, sep=",", row.names="id")
```

Using `library(readr)` (*.csv*):
```{r}
data <- read_csv("filename.csv")
```

### Excel (.xls & .xlsx)

Using `library(readxl)` (*.xls* and *.xlsx*):
```{r}
data <- read_excel("filename.xlsx")
data <- read_excel("filename.xlsx", sheet = "sheet1", col_names = F, col_types = NULL, skip = 2)
data <- 
  read_excel("iqview_blocs_sectors.xls",
             range = "A4:B4663",
             col_names = c("Col1", "Col2"),
             col_types = c("numeric", "guess")
  )
```

Using `library(xlsx)` (*.xls* and *.xlsx*):
```{r}
file <- system.file("tests", "filename.xlsx", package = "xlsx")
# Preserves the type of data
read.xlsx(file, sheetIndex, header=TRUE, colClasses=NA)
# Faster
read.xlsx2(file, sheetIndex, header=TRUE, colClasses="character")
```

Using `library(openxlsx)` (*.xslx*):
```{r}
read.xlsx("filename.xlsx", colNames = F, rows = 1:2, cols = 2:4, detectDates = TRUE)
```



## Export Data

### CSV

Base R:
```{r}
write.table(data, file = "MyData.csv", row.names=FALSE, na="", col.names=FALSE, sep=",")
```

Using *readr*:
```{r}
write_csv(data, "challenge-2.csv")
```

### Excel (.xls & .xslx)

Using *xlsx* (*.xls* and *.xlsx*):
```{r}
write.xlsx(data, "filename.xls", sheetName="Sheet1", col.names=TRUE, row.names=TRUE, append=FALSE)
write.xlsx(data, "filename.xlsx")
# faster
write.xlsx2(data, "filename.xlsx")
```

Using *openxlsx* (*.xlsx*):
```{r}
# basic
write.xlsx(data, file = "export.xlsx", colWidths = "auto")

# with custom style
hs1 <- createStyle(fgFill = "#e96b0a", halign = "CENTER", border = "Bottom", fontColour = "white")
write.xlsx(data,
  file = "export.xlsx",
  colWidths = "auto",
  borders = "surrounding",
  headerStyle = hs1
)

# as a tabel
write.xlsx(data,
  file = "export.xlsx",
  colWidths = "auto",
  firstActiveRow = 2, # freeze the first row
  firstActiveCol = 3, # freeze 2 colums
  asTable = TRUE, # export as a table (style below)
  tableStyle = "TableStyleLight8",
)
```


#### Excel Workbook

You can create a workbook using `library(openxlsx)`.

```{r}
# First have to create a *workbook* using `create_wb()`.
wb <- createWorkbook(title = "DataExport", creator = "TheCreator")

# Then you have to create a new sheet using `wb_sheet()`. You can modify a row using `wd_row()`.
addWorksheet(wb, "NewSheet")
writeData(wb, "NewSheet", data) 

# Finally you can create the Excel file in the working folder using `save_wb()`:
file_name <- "filename.xlsx"
local_path <- dirname(rstudioapi::getSourceEditorContext()$path)
saveWorkbook(wb, paste(local_path, file_name, sep="/"), TRUE)
```

One chunk export:
```{r}
file_name <- "BigAnalysis"
wb <- createWorkbook(title = file_name)
addWorksheet(wb, "Data")
writeData(wb, "Data", unnest)
file_ext <- ".xlsx"
local_path <- dirname(rstudioapi::getSourceEditorContext()$path)
saveWorkbook(wb, paste(local_path, paste(file_name, file_ext, sep=""), sep="/"), TRUE)
```


## Other

- LaTeX array generator: <https://www.tablesgenerator.com/#>
  - Add \begin{table}[ht]
  
Table and lists generation package:
```{r}
install.packages('pander')
```


## Nested Operations

```{r}
# create a nested data frame
by_country <- gapminder %>%
  group_by(country, continent) %>%
  nest()

# apply a function to all the nested lists
custom_function <- function(df) lm(lifeExp ~ year, data = df)
new_col <- map(by_country$data, custom_function)

# merge the new_col
by_country <- by_country %>%
  mutate(model = map(data, new_col))

# compute the residuals for each row
by_country <- by_country %>%
  mutate(resids = map2(data, model, add_residuals))

# expands for each sublist data row and resid
output <- unnest(by_country, resids)
```

